Exploring the conformational dynamics of a large molecule in solution using an external electric field: A Nonequilibrium Molecular Dynamics Simulation

### 2013-09-04 writedown/npt-neq/alanine.tex
revision due to Christof

In this perspective, if one wants to use MD, the proper way to do it would be that of treating a situation of non equilibrium with algorithms that can explicitly describe it. This is the approach that will be followed in this work.

The method we employ is the dynamical nonequilibrium molecular dynamics simulation (D-NEMD).
The method was firstly developed by one of us (GC) and coworkers~\cite{ciccotti1975direct, ciccotti1979thought}.
It has been recently applied, with success, to the study of
hydrodynamics~\cite{orlandini2011hydrodynamics,
  orlandini2011hydrodynamics-01}. The essential feature of D-NEMD is the following: It provides a way of evaluating
nonequilibrium observables by using a properly determined ensemble of nonequilibrium MD
simulations.  In this context, the method can be used for analyzing the
conformational changes of a solvated molecule, subject to a time-dependent EF.
In the present paper, we study a small peptide, alanine dipeptide, as a model for showing how the method can be used. 
The choice of a small peptide, rather than a larger one, is done firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding a possible interplay between
different peptide segments along a chain.
Secondly, the computational cost required is rather massive and thus the testing of the various methodology developments proposed here would not be easy to accomplish for large molecules. Of course, given a positive coming out for such a model system, the aim is that of approaching systems of growing complexity.
The paper is organized as follows:
We firstly describe the idea of D-NEMD, then
introduce the physical
observables for analyzing the nonequilibrium properties. The numerical results of alanine
dipeptide under static and sine-shaped oscillatory EF are
finally presented. By calculating the nonequilibrium properties from D-NEMD
simulations, we are able to address the question of how the peptide
changes the conformation in a EF (conformational space), when do these changes happens (time scale),
and how likely they are (probability). 
To our knowledge, this study is one of the few (if not the first)  where non equilibrium is explicitly considered for a large molecule in solution subject to time-dependent EF.

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the approach for performing dynamical nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known and is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that he dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x)$, where $\mathcal H$ is the Hamiltonian, and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = e^{-iL(t)} \rho(\vect x, 0)$.
On the other hand
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t))
\end{align}
This equation can be formally solved by $\hat O(\vect x(t)) = e^{iL(t)} \hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), e^{-iL(t)} \rho(\vect x, 0)\rangle
  = \langle e^{iL(t)}\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle
\end{align}
Since we assume that the system starts from the distribution of equilibrium (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable measured along trajectories, the
initial configurations of which are subject to the \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the Hamiltonian dynamics is integrated until
time $t$. These trajectories are also called \emph{branching
  trajectories} in this paper (see Fig.\ref{branching} for a pictorial representation). 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.

@@

% The aforementioned D-NEMD simulation algorithm
% does not describe how the boundary of the system should be set up.
% The very naive idea of doing the simulation
% is to simulate a infinitely large isolated system, or any large enough
% isolated subsystem of it.
% In practice, 
% Therefore, we are interested in a 
For this study We are interested in 
nonthermal~\cite{delaHoz2005microwaves} effects of the EF
on the configuration of the solvated alanine dipeptide.
For this reason a proper control of the temperature in a nonequilibrium system
becomes an important issue; in fact it was argued that  an
accurate control of the internal reaction temperature is essential for being able to
perform reproducible experiments~\cite{damm2012can}.
A straightforward (and realistic) implementation of the temperature control consists in 
embedding the alanine into an infinitely large solvent environment, and 
the EF is only applied to the neighborhood of the alanine. In this case
any extra heat generated by the EF can be effectively
dispersed in the large solvent environment.
In this way one can identify the effect of the EF on the conformational space of the solvated molecule,
without the artifact that the surrounding solvent cannot release the additional energy of the EF.
However, in practice, in most of the cases the computational cost of having large solvent environment is rather massive if not prohibitive.
Therefore, it is of interest to substitute the infinitely large
solvent environment with a smaller affordable system and at the same time accurately reproduce the nonequilibrium
physical properties of interest.

As it is well known, in MD, for equilibrium situation, the solution is to couple
the system to a an external thermostat, under which the desired ensemble
(canonical in this case) can be
sampled by simulating a finite size system with periodic boundary
conditions.
However, coupling the system to a thermostat for non equilibrium situations is a delicate issue.
In fact, since a thermostat is explicitly designed to bring the system at equilibrium, any perturbation produced by the EF, would be mostly adsorbed by the thermostat. This means that the physics of the original Hamiltonian dynamics with the EF would be lost and the response to the perturbation becomes unphysical. 
%will disturb the original Hamiltonian dynamics. This is
%in general problematic, 
%because the dynamics  under thermostating is artificial designed,
%so that the equilibrium distribution can be correctly produce, but not
%the nonequilibrium properties of system.
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% However, for a nonequilibrium system, the \emph{dynamics} properties
% will be distrubed by the thermostat, because the
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% For a nonequilibrium simulation, simply couple the system
% to a thermostat may not be a good idea, because the \emph{dynamics}
% generated by the thermostat is artificial, and deviation from the
% Hamiltonian dynamics under investigation will introduce artificial
% a effects in the nonequilibrium observations.
In this section we propose a procedure to solve the problem: we
observe the Hamiltonian dynamics (without thermostat) of the alanine 
itself and the water molecules in the closer solvation shells is
crucial in order to determine the response of the observable of interest
under the perturbation of the EF (in this work, it is the conformational change of alanine).
Instead the detailed dynamics of
the water molecules far away is not of primary importance, that is the region beyond the first solvation shells plays the role of a thermodynamic bath only. 
Therefore, instead of simulating an infinitely
large system, we only simulate a finite size system with periodic
boundary condition with the following characteristics. The simulating region is divided in two
subregions: a spherically defined region around the alanine dipeptide, centered at the
alpha-carbon of radius $R_{ex}$. Here the dynamics is
Hamiltonian and molecules are not subject to a thermostat; we refer to such subregion as \emph{the Hamiltonian dynamics
region}. Instead beyond $R_{ex}$, the dynamics of water is coupled
to a Langevin thermostat, and this region is called \emph{the thermostated
region} (see Fig.~\ref{fig:tmp2}). For technical convenience we
fix the position of the alpha-carbon in space, so that the alanine is always
located at the center of the simulation region.
Therefore, a realistic non equilibrium  dynamics is preserved in the Hamiltonian dynamics region (where the properties of interest are observed)
and possible direct artificial effects of the thermostating process in the region outside are negligible because of the finite correlation
range of liquid water.
At the same time, the thermostated region works as a infinitely large
environment that effectively absorbs the extra heat produced by the EF.
The validity of the above
statement will be later on carefully checked by a series of numerical tests where it is shown that
the response of observables in nonequilibrium (i.e. under the effect of the EF) does not depend on the size of the system
and on the size of the Hamiltonian dynamics subregion, provided that they are
reasonably large.

@@

\subsection{System settings and simulation protocol}
At time $t=0$~ps, the system
has been fully equilibriated without any EF. From $t=0$ to
warm-up time $t=t_{\textrm{warm}}$, the eclectic field is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction, in fact the dipole of the equilibrium configurations of
alanine can be along any direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, the constant EF is
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ as $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
as $t \geq t_{warm}$.
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and solved in 644 TIP3P
water molecules.
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  Hamiltonian dynamics region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
First, an equilibrium NVT simulation of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
% They server as starting point of the branching
% trajectories.
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by Gromacs.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
The whole system is also coupled to a Parrinello-Rahman barostat with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at the ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizeable effect on the
dynamics of the system.

@@

For each branching trajectory,  the molecular configuration of alanine is projected on
the Ramachandran histogram, as
shown in Fig.~\ref{fig:tmp4}. The equilibrium distribution of the
conformations is given in the plot (a), and the fully relaxed
conformation under the action of $E_{\infty} = 1$~V/nm is given  plot (b).  It is
evident from the plot that the conformations are grouped in several clusters, and the
position of the clusters does not change when the external electric
field is applied.
% {\bf LUIGI COMMENT: HAN below is really not clear}\\
% This means that only few of the total of conformations are likely to be visited,
% in the nonequilibrium process.
% These conformations are called ``metastable states''.
% The strict definition of the
% metastability is not exactly the same as those conformations that are likely to be observed.
% Since in this paper we do not focus on the discussion of the definition for the metastability of the system,
% we do not distinguish the definations, and assume they have the same meaning.
% {\bf END COMMENT}\\
For the sake of clarity, we divide the
Ramachandran histogram into 5 subregions, see Fig.~\ref{fig:tmp4},
then the analysis of conformations on the Ramachandran histogram is reduced to the
analysis of the 5 ranges of conformations, i.e. $\{A_1, A_2, B_1, B_2, C\}$,
or in short we call them conformations (the precise meaning is clear
from the context).
In this paper we use:
$A_1 = \{(\phi, \psi) | -134 \leq \phi <  13, -125 \leq \psi < 74\}$,
$A_2 = \{(\phi, \psi) |  128 \leq \phi <-134, -125 \leq \psi < 74\}$,
$B_1 = \{(\phi, \psi) | -110 \leq \phi <  13,   74 \leq \psi <-125\}$,
$B_2 = \{(\phi, \psi) |  128 \leq \phi <-110,   74 \leq \psi <-125\}$,
$C   = \{(\phi, \psi) |   13 \leq \phi < 128, -180 \leq \psi < 180\}$.
Considering the periodic bondary condition, if we write, for example,
$128 \leq \phi <-134$, it means that $\phi$ is in either range
$128 \leq \phi < 180$ or in $-180 \leq \phi < -134$.
These conformations
correspond to different secondary structures of a peptide chain:
$A_1$ and $A_2$ are
corresponding to the alpha helix. $B_1$ and $B_2$ are
corresponding to the beta sheet. $C$ is corresponding to the
left-handed alpha helix.

From Fig.~\ref{fig:tmp4}, it is obvious that the population in each
range of conformations changes when the external EF is applied.
For instance, the population of conformation $C$ vanishes in equilibrium,
but plays an important role when the EF is applied. We
therefore want to study how the alanine changes from one conformation
to the other, how fast is the change, and the relations to the 
the EF. 
First, we calculate the probability of being in a certain conformation range,
defined as:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
% which is the time dependent probability of the alanine being in metastable
% state $I$.
% Secondly, we study the
% forward and backward transition probability of the metastable conformations, which is defined by
% \begin{align}
%   P^{\fwd}_{J,I}(t) & = \mathbb P( \vect x_{t+\Delta t} \in J | \vect x_t \in I) \\
%   P^{\bwd}_{J,I}(t) & = \mathbb P( \vect x_{t-\Delta t} \in J | \vect x_t \in I) 
%   % \qquad I, J \in \{A_1, A_2, B_1, B_2, C\},
% \end{align}
% where $\Delta t$ is the time-scale of the observation.
% For a revserible and homogeneous Markov process, it is easy to show
% that $P^{\fwd}_{J,I} = P^{\bwd}_{J,I}$, and both of them are time independent. For a general nonequilibrium
% process considered by the present work, this relation is genarlly not guarenteed.
Second,
we study the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & =\lim_{\Delta t\rightarrow 0} \frac1{\Delta t} [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
$F_{J,I}(t)$ is essentially the rate of how much conformation $J$ changes
into $I$ at time $t$. A positive value indicates the
net flux from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the original overall nonequilibrium process $\vect x_t$
is reduced to the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
If the resulting reduced process 
is time-reversible, then it is obvious that
$ F_{J,I}(t) = 0$. In an other word, the quantity $ F_{J,I}(t) $ represents
the irreversibility of the reduced process on
$\{A_1, A_2, B_1, B_2, C\}$.
In practice, to measure $ F_{J,I}(t)$, one need first to approximate the limit
in Eq.~\eqref{eqn:tmp7} by the finite difference,
then calculate the joint probabilities on
the right-hand-side from the molecular trajectories.
In this context, for a precise finite difference approximation a smaller $\Delta t$ is
preferable although the statistical uncertainty of measuring the joint probabilities
is higher. In practice,
we use a $\Delta t$ that is small enough for a precise finite difference approximation,
and not small enough so that the joint probabilities can be calculated with acceptable accuracy.
In the case of oscillatory EF, the net probability flux is also highly oscillatory,
thus it is convenient to study the integrated probability flux:
\begin{align}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
this expresses the average over time of the effects of the EF on the conformation of the molecule.

>>>>

In this perspective, if one wants to use MD, the proper way to do it would be that of treating a situation of nonequilibrium with algorithms that can explicitly describe it. This is the approach that will be followed in this work.

The method we employ is the dynamical nonequilibrium molecular dynamics simulation (D-NEMD).
The method was firstly developed by one of us (GC) and coworkers~\cite{ciccotti1975direct, ciccotti1979thought}.
It has been recently applied, with success, to the study of
hydrodynamics~\cite{orlandini2011hydrodynamics,
  orlandini2011hydrodynamics-01}. The essential feature of D-NEMD is the following: It provides a way of evaluating
nonequilibrium observables by using a properly determined ensemble of nonequilibrium MD
simulations.  In this context, the method can be used for analyzing the
conformational changes of a solvated molecule, subject to a time-dependent EF.
In the present paper, we study a small peptide, alanine dipeptide, as a model for showing how the method can be used. 
The choice of a small peptide, rather than a larger one, is done firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding a possible interplay between
different peptide segments along a chain.
Secondly, the computational cost required is rather massive and thus the testing of the various methodology developments proposed here would not be easy to accomplish for large molecules. Of course, given a positive outcome for such a model system, the aim is that of approaching systems of growing complexity.
The paper is organized as follows:
We firstly describe the idea of D-NEMD, then
introduce the physical
observables for analyzing the nonequilibrium properties. The numerical results of alanine
dipeptide under static and sine-shaped oscillatory EF are
finally presented. By calculating the nonequilibrium properties from D-NEMD
simulations, we are able to address the question of how the peptide
changes the conformation in an EF (conformational space), when do these changes happens (time scale),
and how likely they are (probability). 
To our knowledge, this study is one of the few (if not the first)  where nonequilibrium w.r.t. conformation is explicitly considered for a large molecule in solution subject to time-dependent EF.

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the approach for performing dynamical nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known and is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that he dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x)$, where $\mathcal H$ is the Hamiltonian, and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = e^{-iL(t)} \rho(\vect x, 0)$.
On the other hand
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t))
\end{align}
This equation can be formally solved by $\hat O(\vect x(t)) = e^{iL(t)} \hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), e^{-iL(t)} \rho(\vect x, 0)\rangle
  = \langle e^{iL(t)}\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle
\end{align}
Since we assume that the system starts from the distribution of equilibrium (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable measured along trajectories, the
initial configurations of which are subject to the \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the Hamiltonian dynamics is integrated until
time $t$. These trajectories are also called \emph{branching
  trajectories} in this paper (see Fig.\ref{branching} for a pictorial representation). 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.
Stochastic dynamics (Langevin) can be hadeled analogoursly.

@@

% The aforementioned D-NEMD simulation algorithm
% does not describe how the boundary of the system should be set up.
% The very naive idea of doing the simulation
% is to simulate a infinitely large isolated system, or any large enough
% isolated subsystem of it.
% In practice, 
% Therefore, we are interested in a 
For this study, we are interested in 
nonthermal~\cite{delaHoz2005microwaves} effects of the EF
on the configuration of the solvated alanine dipeptide.
For this reason a proper control of the temperature in a nonequilibrium system
becomes an important issue; in fact it was argued that  an
accurate control of the internal reaction temperature is essential for being able to
perform reproducible experiments~\cite{damm2012can}.
A straightforward (and realistic) implementation of the temperature control consists in 
embedding the alanine into an infinitely large solvent environment, and 
the EF is only applied to the neighborhood of the alanine. In this case
any extra heat generated by the EF can be effectively
dispersed in the large solvent environment.
In this way one can identify the effect of the EF on the conformational space of the solvated molecule,
without the artifact that the surrounding solvent cannot release the additional energy of the EF.
However, in practice, in most of the cases the computational cost of having large solvent environment is rather massive if not prohibitive.
Therefore, it is of interest to substitute the infinitely large
solvent environment with a smaller affordable system and at the same time accurately reproduce the nonequilibrium
physical properties of interest.

As it is well known, in MD, for equilibrium situations, the solution is to couple
the system to a an external thermostat, under which the desired ensemble
(canonical in this case) can be
sampled by simulating a finite size system with periodic boundary
conditions.
However, coupling the system to a thermostat for nonequilibrium situations is a delicate issue.
In fact, since a thermostat is explicitly designed to bring the system into equilibrium, any perturbation produced by the EF would be mostly adsorbed by the thermostat. This means that the physics of the original Hamiltonian dynamics with the EF would be lost and the response to the perturbation becomes unphysical. 
%will disturb the original Hamiltonian dynamics. This is
%in general problematic, 
%because the dynamics  under thermostating is artificial designed,
%so that the equilibrium distribution can be correctly produce, but not
%the nonequilibrium properties of system.
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% However, for a nonequilibrium system, the \emph{dynamics} properties
% will be distrubed by the thermostat, because the
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% For a nonequilibrium simulation, simply couple the system
% to a thermostat may not be a good idea, because the \emph{dynamics}
% generated by the thermostat is artificial, and deviation from the
% Hamiltonian dynamics under investigation will introduce artificial
% a effects in the nonequilibrium observations.
In this section we propose a procedure to solve the problem: we
observe the Hamiltonian dynamics (without thermostat) of the alanine 
itself and the water molecules in the closer solvation shells is
crucial, in order to determine the response of the observable of interest
under the perturbation of the EF (in this work, it is the conformational change of alanine).
Instead the detailed dynamics of
the water molecules far away is not of primary importance, that is the region beyond the first solvation shells plays the role of a thermodynamic bath only. 
Therefore, instead of simulating an infinitely
large system, we only simulate a finite size system with periodic
boundary condition with the following characteristics. The simulating region is divided in two
subregions: a spherically defined region around the alanine dipeptide, centered at the
alpha-carbon of radius $R_{ex}$. Here the dynamics is
Hamiltonian and molecules are not subject to a thermostat; we refer to such subregion as \emph{the Hamiltonian dynamics
region}. Instead beyond $R_{ex}$, the dynamics of water is coupled
to a Langevin thermostat, and this region is called \emph{the thermostated
region} (see Fig.~\ref{fig:tmp2}). For technical convenience we
fix the position of the alpha-carbon in space, so that the alanine is always
located at the center of the simulation region.
Therefore, a realistic nonequilibrium  dynamics is preserved in the Hamiltonian dynamics region (where the properties of interest are observed)
and possible direct artificial effects of the thermostating process in the region outside are negligible because of the finite correlation
range of liquid water.
At the same time, the thermostated region works as a infinitely large
environment that effectively absorbs the extra heat produced by the EF.
The validity of the above
statement will be later on carefully checked by a series of numerical tests where it is shown that
the response of observables in nonequilibrium (i.e. under the effect of the EF) does not depend on the size of the system
and on the size of the Hamiltonian dynamics subregion, provided that they are
reasonably large.

@@

\subsection{System settings and simulation protocol}
At time $t=0$~ps, the system
has been fully equilibriated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the eclectic field is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction, in fact the dipole of the equilibrium configurations of
alanine can be along any direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, the constant EF is
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ as $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
as $t \geq t_{warm}$.
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and solved in 644 TIP3P
water molecules.
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  Hamiltonian dynamics region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
First, an equilibrium NVT simulation of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
% They server as starting point of the branching
% trajectories.
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by Gromacs.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
The whole system is also coupled to a Parrinello-Rahman barostat with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at the ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizeable effect on the
dynamics of the system.

@@

For each branching trajectory,  the molecular configuration of alanine is projected on
the Ramachandran histogram, as
shown in Fig.~\ref{fig:tmp4}. The equilibrium distribution of the
conformations is given in plot (a), and the fully relaxed
conformation under the action of $E_{\infty} = 1$~V/nm is given in plot (b).  It is
evident from the plot that the conformations are grouped in several clusters, and the
position of the clusters does not change when the external electric
field is applied.
% {\bf LUIGI COMMENT: HAN below is really not clear}\\
% This means that only few of the total of conformations are likely to be visited,
% in the nonequilibrium process.
% These conformations are called ``metastable states''.
% The strict definition of the
% metastability is not exactly the same as those conformations that are likely to be observed.
% Since in this paper we do not focus on the discussion of the definition for the metastability of the system,
% we do not distinguish the definations, and assume they have the same meaning.
% {\bf END COMMENT}\\
For the sake of clarity, we divide the
Ramachandran histogram into 5 subregions, see Fig.~\ref{fig:tmp4},
then the analysis of conformations on the Ramachandran histogram is reduced to the
analysis of the 5 ranges of conformations, i.e. $\{A_1, A_2, B_1, B_2, C\}$,
or in short we call them conformations (the precise meaning is clear
from the context).
In this paper we use:
$A_1 = \{(\phi, \psi) | -134 \leq \phi <  13, -125 \leq \psi < 74\}$,
$A_2 = \{(\phi, \psi) |  128 \leq \phi <-134, -125 \leq \psi < 74\}$,
$B_1 = \{(\phi, \psi) | -110 \leq \phi <  13,   74 \leq \psi <-125\}$,
$B_2 = \{(\phi, \psi) |  128 \leq \phi <-110,   74 \leq \psi <-125\}$,
$C   = \{(\phi, \psi) |   13 \leq \phi < 128, -180 \leq \psi < 180\}$.
Considering the periodic bondary condition, if we write, for example,
$128 \leq \phi <-134$, it means that $\phi$ is in either range
$128 \leq \phi < 180$ or in $-180 \leq \phi < -134$.
These conformations
correspond to different secondary structures of a peptide chain:
$A_1$ and $A_2$ are
corresponding to the alpha helix. $B_1$ and $B_2$ are
corresponding to the beta sheet. $C$ is corresponding to the
left-handed alpha helix.

From Fig.~\ref{fig:tmp4}, it is obvious that the population in each
range of conformations changes when the external EF is applied.
For instance, the population of conformation $C$ vanishes in equilibrium,
but plays an important role when the EF is applied. We
therefore want to study how the alanine changes from one conformation
to the other, how fast the change is, and the relations to the 
the EF. 
First, we calculate the probability of being in a certain conformation range,
defined as:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
% which is the time dependent probability of the alanine being in metastable
% state $I$.
% Secondly, we study the
% forward and backward transition probability of the metastable conformations, which is defined by
% \begin{align}
%   P^{\fwd}_{J,I}(t) & = \mathbb P( \vect x_{t+\Delta t} \in J | \vect x_t \in I) \\
%   P^{\bwd}_{J,I}(t) & = \mathbb P( \vect x_{t-\Delta t} \in J | \vect x_t \in I) 
%   % \qquad I, J \in \{A_1, A_2, B_1, B_2, C\},
% \end{align}
% where $\Delta t$ is the time-scale of the observation.
% For a revserible and homogeneous Markov process, it is easy to show
% that $P^{\fwd}_{J,I} = P^{\bwd}_{J,I}$, and both of them are time independent. For a general nonequilibrium
% process considered by the present work, this relation is genarlly not guarenteed.
Second,
we study the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\, [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
Here the time interval $\Delta t$ should be small enough comparing with the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be linearly approximated.
In the meanwhile,  $\Delta t$ should also be  large enough comparing with
the timestep of MD integrator, 
so that the left hand of Eq.~\eqref{eqn:tmp7} is well-defined, and can be
estimated with good statistical accuracy.
See also the discussions in Ref.~\cite{schuette2011markov}.
$F_{J,I}(t)$ is essentially the rate of how much conformation $J$ changes
into $I$ at time $t$. A positive value indicates the
net flux from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the original overall nonequilibrium process $\vect x_t$
is reduced to the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
If the resulting reduced process 
is time-reversible, then it is obvious that
$ F_{J,I}(t) = 0$. In an other word, the quantity $ F_{J,I}(t) $ represents
the irreversibility of the reduced process on
$\{A_1, A_2, B_1, B_2, C\}$.
% In practice, to measure $ F_{J,I}(t)$, one needs first to approximate the limit
% in Eq.~\eqref{eqn:tmp7} by the finite difference,
% then calculate the joint probabilities on
% the right-hand-side from the molecular trajectories.
% In this context, for a precise finite difference approximation a smaller $\Delta t$ is
% preferable although the statistical uncertainty of measuring the joint probabilities
% is higher. In practice,
% we use a $\Delta t$ that is small enough for a precise finite difference approximation,
% and not small enough so that the joint probabilities can be calculated with acceptable accuracy.
In the case of oscillatory EF, the net probability flux is also highly oscillatory,
thus it is convenient to study the integrated probability flux:
\begin{align}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
this expresses the average over time of the effects of the EF on the conformation of the molecule.
